---
title: "CHR Data"
author: "Gabriel Kahn"
date: "May 5, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(stringr)
CHR_data <- read_csv("CHR_data.csv")

```
# Introduction

This report is intended to be the first step in a larger project analyzing how health and economic factors influence US presidential elections. Here, I will use data collected from County Health Rankings and Roadmaps, an organization formed as a partnership between the Robert Wood Johnson Foundation and the University of Wisconsin Population Health Institute. The purpose here is exploratory; the hope is to gain a more nuanced look at the nation's health at a county level, with a particular focus on swing states. Looking ahead, I hope to collect data on election results by county to see if health factors relate to election outcomes.


# Tidying Data

```{r}
CHR_data <- CHR_data[-c(1),]
# remove the redundant first row
CHR_data <-  CHR_data %>%
   data.frame(lapply(CHR_data, trimws), stringsAsFactors = FALSE) 
# Take spaces out of column names
CHR_data_num <- CHR_data                           # Replicate data
CHR_data_num[ , 8:100] <- as.data.frame(   # Recode characters as numeric
  apply(CHR_data_num[ , 8:100], 2, as.numeric))

CHR_data_num                 
```

```{r}

CHR_data %>%

  
